Fan, splint and branching rules.

### 2011-10-13 AffineLieAlgebras/Splints/paper.tex
Paper almost finished

\section{Introduction}
\label{sec:Introduction} Splint $\phi$ of a root system $\Delta_1$
into a root system $\Delta$ is a bijective map of roots of
$\Delta_{1}$ to (proper) subset of $\Delta$ which commutes with
vector composition law in $\Delta_{1}$ and $\Delta$.
\begin{equation*}
\phi:\Delta_1 \longrightarrow \Delta
\end{equation*}
\begin{equation*}
\phi \circ (\alpha + \beta) =\phi \circ \alpha + \phi \circ \beta,
\,\,\, \alpha,\beta \in \Delta_1
\end{equation*}

@@

Consider a simple Lie algebra $\mathfrak{g}$ and its regular subalgebra $%
\mathfrak{a}\longrightarrow \mathfrak{g}$ such that $\mathfrak{a}$ is a
reductive subalgebra $\mathfrak{a
\subset g}$ with correlated root spaces: $\mathfrak{h}_{\mathfrak{a}}^{\ast
}\subset \mathfrak{h}_{\mathfrak{g }}^{\ast }$. Let $\mathfrak{a}^S$ be a
semisimple summand of $\mathfrak{a}$, this means that $\mathfrak{a}=%
\mathfrak{a}^S \oplus \mathfrak{u}(1)\oplus \mathfrak{u}(1)\oplus \dots$. We
shall consider $\mathfrak{a}^S$ to be a proper regular subalgebra and $%
\mathfrak{a}$ to be the maximal subalgebra with $\mathfrak{a}^S$ fixed that
is the rank $r$ of $\frak{a}$ is equal to that of $\mathfrak{g}$.

@@

In this Section we consider the situation where the splinting subalgebra is
''larger'' than the Lie algebra $\frak{s}$ defined by the root system $%
\Delta _{\frak{s}\,0}=CoIm_{\phi }\left( \Delta _{\frak{s}}\right) $ (the
details will be given below). In such a case to find branching coefficients
for a splinting injection $\frak{a}\hookrightarrow \frak{g}$ means to find
weight multiplicities for an irreducible $\frak{s}$-module $L_{\frak{s}%
}^{\nu }$ with fixed highest weight $\nu $. Notice that $\frak{s}$ must not
be a subalgebra of $\frak{g}$.

@@

\label{sec:conclusions}

\section*{Acknowledgements}
\label{sec:acknowledgements}

The work of A.A.~Nazarov is supported by the Chebyshev Laboratory
(Department of Mathematics and Mechanics, Saint-Petersburg State
University) under the grant 11.G34.31.0026 of the Government of the
Russian Federation.

@@

\section{Appendix}

@@

Let us demonstrate that the injection splints
 \begin{figure}[h!bt]
  \noindent\centering{
   \includegraphics[width=100mm]{g2-roots}
  }
  \caption{Weyl group orbit (dotted) producing singular element of $A_{2}$-module $L^{(3,2)}$ and its decomposition into the sum of splint images of singular elements of $A_{1}\oplus A_{1}$-modules (dashed). Weight multiplicities of $A_{1}\oplus A_{1}$-module coincide with branching coefficients for the reduction $L^{(3,2)}_{A_{2}\downarrow A_{1}\oplus u(1)}$.}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

>>>>

\section{Introduction}
\label{sec:Introduction}

Splint $\phi$ of a root system $\Delta_1$ into a root system
$\Delta$ is a bijective map of roots of $\Delta_{1}$ to (proper)
subset of $\Delta$ which commutes with vector composition law in
$\Delta_{1}$ and $\Delta$.
\begin{equation*}
\phi:\Delta_1 \longrightarrow \Delta
\end{equation*}
\begin{equation*}
\phi \circ (\alpha + \beta) =\phi \circ \alpha + \phi \circ \beta,
\,\,\, \alpha,\beta \in \Delta_1
\end{equation*}

@@

Consider a simple Lie algebra $\mathfrak{g}$ and its regular subalgebra $%
\mathfrak{a}\longrightarrow \mathfrak{g}$ such that $\mathfrak{a}$
is a
reductive subalgebra $\mathfrak{a \subset g}$ with correlated root spaces: $%
\mathfrak{h}_{\mathfrak{a}}^{\ast }\subset \mathfrak{h}_{\mathfrak{g }%
}^{\ast }$. Let $\mathfrak{a}^S$ be a semisimple summand of
$\mathfrak{a}$,
this means that $\mathfrak{a}=\mathfrak{a}^S \oplus \mathfrak{u}(1)\oplus %
\mathfrak{u}(1)\oplus \dots$. We shall consider $\mathfrak{a}^S$
to be a proper regular subalgebra and $\mathfrak{a}$ to be the
maximal subalgebra with $\mathfrak{a}^S$ fixed that is the rank
$r$ of $\frak{a}$ is equal to that of $\mathfrak{g}$.

@@

In this Section we study properties of injective splints. It will
be demonstrated that provided such splint exists to find branching
coefficients for a splinting injection $\frak{a}\hookrightarrow
\frak{g}$ means to find weight multiplicities of an irreducible
$\frak{s}$-module $L_{\frak{s}}^{\nu }$ with fixed highest weight
$\nu $. Notice that $\frak{s}$ must not be a subalgebra of
$\frak{g}$.

@@

\label{sec:conclusions}It is explicitly demonstrated that splint
presents a very effective tool to find branching coefficients. In
particular injective splints provide a possibility to reduce
branching rules calculations for highest weight modules to a
determination of weight multiplicities for a module with the same
Dynkin labels referred to a different Lie algebra. This algebra
$\frak{s}$ must not be a subalgebra in the initial $\frak{g}$ , it
has the same rank $r_{\frak{s}}=r$\ , but as a rule is less
complicated than $\frak{g}$ .

It is significant that for the injections $D_{r}\hookrightarrow B_{r}$ , $%
D_{r}\hookrightarrow C_{r}$ and $A_{r}\hookrightarrow
A_{r-1}\oplus u\left( 1\right) $ splint technique shows
immediately Gelfand-Tzeytlin rules for branching: the nonzero
branching coefficients are equal to 1, the reduction is
multiplicity free. Here it is an immediate consequence of the
structure of the second stem being a direct sum of $A_{1}$
algebras and the fact that the corresponding module
$L_{\frak{s}}^{\mu }$ is irreducible.

\section*{Acknowledgements}
\label{sec:acknowledgements}
The work was supported in part by the
RFFI grant N 09-01-00504. The work of A.A.~Nazarov is supported by
the Chebyshev Laboratory (Department of Mathematics and Mechanics,
Saint-Petersburg State University) under the grant 11.G34.31.0026
of the Government of the Russian Federation.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section*{Appendix}

@@

Let us demonstrate that for injective splints of classical Lie
algebras the following property is valid:

Let $\Delta \approx (\Delta _{\frak{a}},\Delta _{\frak{s}})$ be an
injective
splint with the decomposition of simple roots $S=S_{\frak{c}}\cup S_{\frak{d}%
}$ with $S_{\frak{c}}=S\cap S_{\frak{a}}$ and $S_{\frak{d}}=S\cap S_{\frak{s}%
}$.

Then for any simple root $\beta \in S_{\frak{s}}$ there exists the
pair of
roots ( $\alpha $ ,$\beta ^{\prime }$) with $\alpha \in $ $S_{\frak{c}%
},\beta ^{\prime }\in S_{\frak{s}}$ such that $\alpha =\beta
-\beta ^{\prime }$

Type 1. $\Delta _{G_{2}}\approx (\Delta _{A_{2}},\Delta
_{A_{2}}).$

 \begin{figure}[h!bt]
  \noindent\centering{
   \includegraphics[width=100mm]{g2-roots}
  }
  \caption{Weyl group orbit (dotted) producing singular element of $A_{2}$-module $L^{(3,2)}$ and its decomposition into the sum of splint images of singular elements of $A_{1}\oplus A_{1}$-modules (dashed). Weight multiplicities of $A_{1}\oplus A_{1}$-module coincide with branching coefficients for the reduction $L^{(3,2)}_{A_{2}\downarrow A_{1}\oplus u(1)}$.}
\end{figure}
Here both stems are metric and the corresponding root systems are
equivalent. In Figure 4 a part of the singular element $\Phi
_{G_{2}}^{\left( 0\right) }$ is presented. The boundaries of $\bar{C_{\frak{a%
}}}$ are the dashed lines starting in the center of the singular
element. It
contains the edge $\lambda _{2}=-\alpha _{2}=-\beta _{2}$ and the roots $%
-\beta _{1}=-s_{\alpha _{2}}\circ \beta _{3}$ , $-\beta _{3}$
($\beta _{3}$ is indicated as $\beta _{l}$). For this root $\beta
_{1}$ the necessary pair
is ( $\alpha _{1}$ ,$\beta _{2}$): $\alpha _{1}=\beta _{1}-\beta _{2}$. The $%
\lambda _{2,3}^{\frak{s}}=\beta _{3}$ edge is equal to $\lambda _{1}^{\frak{s%
}}=\beta _{1}=s_{\alpha _{2}}\circ \beta _{3}$ and $m_{1}$ indice
is aquired by the $\frak{s}$-module that inherit the second
indice. In this particular
case they are $m_{1}=m_{2}=0$. The general case with the initial module $%
L^{\mu }$ and $\mu =m_{1}\omega _{1}+m_{2}\omega _{2}$.can be
treated in the same way: one finds an edge $\lambda _{2}=-\left(
m_{2}+1\right) \beta _{2}$ and put $\lambda
_{1}^{\frak{s}}=-\left( m_{1}+1\right) \beta _{1}$, its end
belongs to the boundary $\bar{C_{\frak{a}}}$ . The reflection
$s_{\beta
_{2}} $ sends $\beta _{1}$ to $\beta _{3}$ and the corresponding edge $%
\lambda _{2,3}^{\frak{s}}=-\left( m_{1}+1\right) \beta _{3}$ has
the same length. Now consider $\lambda _{1}^{\frak{s}}$ (or
$\lambda _{2,3}^{\frak{s}} $) and $\lambda _{1,3}^{\frak{s}}$ (or
$\lambda _{2,3,1}^{\frak{s}}$) edges to find that they belong to
the boundary $\bar{C_{\frak{a}}}$ and the Weyl symmetry predicts
that $\lambda _{1,3}^{\frak{s}}=-\left( m_{2}+1\right)
\beta _{3}$ ($\lambda _{2,3,1}^{\frak{s}}=-\left( m_{2}+1\right) \beta _{1}$%
) . Finally the edge $\lambda _{1,3,2}^{\frak{s}}=-\left(
m_{1}+1\right) \beta _{2}$ close the polytop . The latter
corresponds to the singular
element $\Phi _{\frak{s}}^{\left( \widetilde{\mu }\right) }=\sum_{w\in W_{%
\frak{s}}}\varepsilon \left( w\right) e^{w\circ \left(
\widetilde{\mu }+\rho
_{\frak{s}}\right) }$ of the module $L_{\frak{s}}^{\left( \widetilde{\mu }%
\right) }$ with $\widetilde{\mu }=m_{1}\widetilde{\omega }_{1}+m_{2}%
\widetilde{\omega }_{2}$ . Notice that in this case the sign
factors can be obtained directly in the initial weight system as
far as the stem is metric.

Type 1. $\Delta _{F_{4}}\approx (\Delta _{D_{4}},\Delta
_{D_{4}}).$

Both stems are metric and the corresponding root systems are
equivalent. The system $\Delta _{D_{4}}$ of the subalgebra
$\frak{a=}D_{4}$ is formed by the set $\left\{ \pm e_{i},\pm
e_{j}\right\} _{|i,j=1,\ldots 4,\quad i\neq j}$ . The simple roots
$S_{\frak{c}}$ are $\left\{ e_{2}-e_{3},e_{3}-e_{4}\right\} $ and
$S_{\frak{d}}=\left\{ e_{4},\frac{1}{2}\left(
e_{1}-e_{2}-e_{3}-e_{4}\right) \right\} $.

For a module $L^{\mu }$ with $\mu =\sum m_{k}\omega _{k}$ consider the edge $%
\lambda _{3}=-\left( m_{3}+1\right) e_{4}=-\left( m_{3}+1\right)
\beta _{3}$
\ Compose an edge $\lambda _{2}^{\frak{s}}=-\left( \widetilde{m}%
_{2}+1\right) \beta _{2}$ . Now the necessary pair is $\left(
\alpha
_{2}=e_{3}-e_{4},\beta _{3}\right) $. The intersection of $\lambda _{2}^{%
\frak{s}}$ with the $\alpha _{2}$ -boundary of
$\bar{C_{\frak{a}}}$ fixes
its length to be $\lambda _{2}^{\frak{s}}=-\left( m_{2}+1\right) \beta _{2}$%
\ and the length of the edge $\lambda _{3,2}^{\frak{s}}$ is equal
to that of
$\lambda _{2}^{\frak{s}}$. Next consider the edge $\lambda _{2}^{\frak{s}%
}=-\left( m_{2}+1\right) \beta _{2}$ and the pair $\left( \alpha
_{1}=e_{2}-e_{3},\beta _{1}=e_{2}\right) $ \ The length of the
edge $\lambda _{1}^{\frak{s}}$ becomes equal to $\lambda
_{1}^{\frak{s}}=-\left( m_{1}+1\right) \beta _{1}$ and so on. The
edges along the roots like $\alpha _{4}=\beta
_{4}=\frac{1}{2}\left( e_{1}-e_{2}-e_{3}-e_{4}\right) $ are
treated similarly and finally the singular element $\Phi
_{\frak{s}}^{\left( \widetilde{\mu }\right) }=\sum_{w\in
W_{\frak{s}}}\varepsilon \left( w\right) e^{w\circ \left(
\widetilde{\mu }+\rho _{\frak{s}}\right) }$ of the module
$L_{\frak{s}}^{\left( \widetilde{\mu }\right) }$ with
$\widetilde{\mu }=\sum m_{k}\widetilde{\omega }_{k}$ is formed in
$\bar{C_{\frak{a}}}$.

Type 2. $\Delta _{B_{r}}\approx (\Delta _{D_{r}},\Delta _{\oplus
^{r}A_{1}}). $

Both stems are metric. The injection is fixed by the stem $\Delta
_{D_{r}}$ with simple roots $S_{\frak{a}}=\left\{
e_{1}-e_{2},e_{2}-e_{3},\ldots
,e_{r-1}-e_{r},e_{r-1}+e_{r}\right\} $ . The second stem
corresponds to a
direct sum of algebras $A_{1}$ and has the simple roots $S_{\frak{s}%
}=\left\{ e_{1},e_{2},\ldots ,e_{r-1},e_{r}\right\} $. Consider the edge $%
\lambda _{r}=-\left( m_{r}+1\right) \beta _{r}$ (here $\beta
_{r}=e_{r}$) and .$\lambda _{r-1}=-\left(
\widetilde{m}_{r-1}+1\right) \beta _{r-1}$
attached to it (here $\beta _{r-1}=e_{r-1}$). The corresponding pair is $%
\left( \alpha _{r-1}=e_{r-1}-e_{r},\beta _{r-1}=e_{r-1}\right) $
.\ The intersection condition fixes the second edge to be $\lambda
_{r-1}=-\left( m_{r-1}+1\right) \beta _{r-1}$ , it is orthogonal
to $\beta _{r}$ so the opposite edge has the same length. The
Dynkin indice $m_{r-1}$ now refers
also to the\ simple root $\beta _{r-1}$ . Next consider the obtained edge $%
\lambda _{r-1}=-\left( m_{r-1}+1\right) \beta _{r-1}$ and $\lambda
_{r-2}=-\left( \widetilde{m}_{r-2}+1\right) \beta _{r-2}$ to
obtain the indice $\widetilde{m}_{r-2}=m_{r-2}$ and the edge
$\lambda _{r-2}=-\left( m_{r-2}+1\right) \beta _{r-2}$ and so on
till all the pairs of edges are properly fixed. Finally in
$\bar{C_{D_{r}}}$ the element $\Phi _{\oplus ^{r}A_{1}}^{\left(
\widetilde{\mu }\right) }=\sum_{w\in W_{\oplus
^{r}A_{1}}}\varepsilon \left( w\right) e^{w\circ \left( \widetilde{\mu }+%
\frac{1}{2}\sum e_{k}\right) }$ can be constructed for the module
$L_{\oplus ^{r}A_{1}}^{\left( \widetilde{\mu }\right) }$ with
$\widetilde{\mu }=\sum m_{k}\frac{1}{2}e_{k}$ .

Type 2. $\Delta _{C_{r}}\approx (\Delta _{D_{r}},\Delta _{\oplus
^{r}A_{1}}). $

The situation in this case is eqivalent to the previous one and
the additional edges are constructed similarly.

Type 3 $\Delta _{A_{r}}\approx (\Delta _{A_{r-1}\oplus
u_{1}},\Delta _{\oplus ^{r}A_{1}}).$

Here only the first stem is metric and it fixes the injection with
simple roots $S_{\frak{a}}=\left\{ e_{1}-e_{2},e_{2}-e_{3},\ldots
,e_{r-1}-e_{r}\right\} $ . The second stem corresponding to a direct sum of $%
r$ copies of $A_{1}$ has the simple roots $S_{\frak{s}}=\left\{
e_{1}-e_{r+1},e_{2}-e_{r+1},\ldots ,e_{r}-e_{r+1}\right\} $.
Consider the edge $\lambda _{r}=-\left( m_{r}+1\right) \beta _{r}$
with $\beta
_{r}=e_{r}-e_{r+1}$ and .$\lambda _{r-1}=-\left( \widetilde{m}%
_{r-1}+1\right) \beta _{r-1}$ \ with $\beta
_{r-1}=e_{r-1}-e_{r+1}$ attached to it . Then the corresponding
pair is $\left( \alpha _{r-1}=e_{r-1}-e_{r},\beta
_{r-1}=e_{r-1}-e_{r+1}\right) $ .\ The intersection with the
boundary of $\bar{C_{A_{r-1}}}$ orthogonal to $\alpha _{r-1}$
fixes the second edge to be $\lambda _{r-1}=-\left(
m_{r-1}+1\right) \beta _{r-1}$. The Dynkin indice $m_{r-1}$ is to
be used for the fundamental weight  $\omega _{r-1}$ . The\
reflection $s_{\beta _{r}}$ sends $\lambda _{r-1}=-\left(
m_{r-1}+1\right) \beta _{r-1}$ to $\lambda _{r,r-1}=-\left(
m_{r-1}+1\right) \beta _{r-1}$\ . Next consider the obtained edge
$\lambda _{r-1}=-\left( m_{r-1}+1\right) \beta _{r-1}$ and
$\lambda _{r-2}=-\left( \widetilde{m}_{r-2}+1\right) \beta _{r-2}$
with $\beta _{r-2}=e_{r-2}-e_{r+1} $ to obtain the indice
$\widetilde{m}_{r-2}=m_{r-2}$ and the edge $\lambda _{r-2}=-\left(
m_{r-2}+1\right) \beta _{r-2}$ and so on till all the pairs of
edges are properly fixed. Finally in $\bar{C_{D_{r}}}$ the element
$\Phi _{\oplus ^{r}A_{1}}^{\left( \widetilde{\mu }\right)
}=\sum_{w\in W_{\oplus
^{r}A_{1}}}\varepsilon \left( w\right) e^{w\circ \left( \widetilde{\mu }+%
\widetilde{\rho }\right) }$ can be constructed for the module
$L_{\oplus ^{r}A_{1}}^{\left( \widetilde{\mu }\right) }$ with
$\widetilde{\mu }=\sum m_{k}\beta _{k}$ . The simplest case
$\Delta _{A_{2}}\approx (\Delta _{A_{1}\oplus u_{1}},\Delta
_{A_{1}\oplus A_{1}})$ is presented in Example 4.1 and Figure 1.

Type 3 $\Delta _{B_{2}}\approx (\Delta _{A_{1}},\Delta _{A_{2}}).$

The necessary elements are presented in Example 4.1 and Figure 1 , $%
S_{A\_1}=\left\{ e_{1}-e_{2}\right\} $ , $S_{A\_2}=\left\{
e_{1},e_{2}\right\} $.The edge $\lambda _{\alpha _{2}}=\lambda
_{\beta _{2}}=-\left( m_{2}+1\right) \beta _{2}$ has an edge
$\lambda _{\beta _{1}}=-\left( \widetilde{m}_{1}+1\right) \beta
_{1}$ attached to it. Consider the pair $\left( \alpha
_{1}=e_{1}-e_{2},\beta _{1}=e_{1}\right) $ .The end of the edge
$\lambda _{\beta _{1}}$ is to define a weight
invariant under the reflexion $s_{\alpha _{1}}$ . This defines its length: $%
\lambda _{\beta _{1}}=-\left( m_{1}+1\right) \beta _{1}$ . In the
coimage of
the second stem, that is in the $A_{2}$ root system, the reflection $%
s_{\beta _{2}}$ sends $\lambda _{\beta _{1}}=-\left(
m_{1}+1\right) \beta
_{1}$ to $\lambda _{2,3}$\ ., thus the latter edge has the same length in $%
\beta _{3}=e_{1}+e_{3}$ , we have $\lambda _{2,3}=-\left(
m_{1}+1\right)
\beta _{3}$ with $\beta _{3}=e_{1}+e_{3}$ . The irreducible $\frak{s}$%
-module has the highest weight $\widetilde{\mu }=m_{1}\widetilde{\omega }%
_{1}+m_{2}\widetilde{\omega }_{2}$ . In Figure 1 we see the
details of these relations in a particular case where
$L_{B_{2}}^{\left[ 3,2\right] }$ is reduced to a subalgebra
$A_{1}\oplus u\left( 1\right) $ and the
corresponding highest weights with their multiplicities form the diagram $%
\mathcal{N}_{A\_2}^{\left[ 3,2\right] }$ .

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
